---
title: "Dashboard Prep"
format: html
editor: visual
---

## HW #7: Dashboard Prep

You've looked at different types of data, conducted exploratory data analysis, manipulated data with dplyr and other tidy methods, created maps.... Now it's time to think about your flexdashboard final project!

## Data

What data are you planning to use? Describe the data here and read it in below.

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
library(scales)
library(plotly)
library(dplyr)
```

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

data <- read_csv(here("data/report.csv"))

##My dataset is a comprehensive summary of different violent crime offenses in the US from 1997-2013, organized by state/capitals. 
```

## Variables of interest

Which variables are interesting to you? What kinds of questions will you be asking as you look into this data? What story is your data telling? Please include three questions here. The code chunk below is optional.

The data that is interesting to me are the variables homicides, report_year, agency_jurisdiction, population, and robberies. I want my visualizations to explore the questions of how do time (report_year), population, and place (agency_jurisdiction) effect the violent crime variables. My data shows how violent crime in the US has changed over a 40 year time period and how it varies from population to population, and state to state.

```{r}

```

## Plots

You can present plots as "answers" to your exploratory questions above. You'll need to generate at least three types of plots to put on your dashboard. At least one of them should be interactive using plotly or girafe. Test them out here, include the accompanying questions that they answer, and explain why/how the plot answers that question.

These can look really basic right now. We'll talk more about themes and layers next week.

```{r}
##First I want to clean my data
data <- data %>%
  clean_names()
#Checking to see the type of data is correct to work with.
skim_data <- skimr::skim(data)
```

```{r}
#structure of the dataset
str(data)

summary(data)

unique(data$report_year)          
unique(data$agency_jurisdiction)  


```

```{r}
data$agency_code <- as.factor(data$agency_code)
data$agency_jurisdiction <- as.factor(data$agency_jurisdiction)
data$report_year<- as.factor(data$report_year)
```

```{r}
#ggplot(data, aes(x = report_year, y = agency_jurisdiction)) +
#  geom_point()
```

```{r}
###First Visualization
# Aggregate 
jurisdiction_robbery_data <- data %>%
  group_by(jurisdiction = sub(",.*", "", agency_jurisdiction), report_year) %>%
  summarize(total_robberies = sum(robberies, na.rm = TRUE)) %>%
  ungroup()
##Creating smaller set: highestpopulation
top_jurisdiction <- jurisdiction_robbery_data %>%
  group_by(jurisdiction) %>%
  summarize(total_robberies_all_years = sum(total_robberies, na.rm = TRUE)) %>%
  arrange(desc(total_robberies_all_years)) %>%
  slice(1:10) %>%
  pull(jurisdiction)

# Filter data 
top_jurisdiction_data <- jurisdiction_robbery_data %>%
  filter(jurisdiction %in% top_jurisdiction)

##Need to mutate factor to numeric so that the x axis labels can change to every 5 years for readability purposes
top_jurisdiction_data <- top_jurisdiction_data %>%
  mutate(report_year = as.numeric(as.character(report_year)))  

# line chart
ggplot(top_jurisdiction_data, aes(x = report_year, y = total_robberies, color = jurisdiction, group = jurisdiction)) +
  geom_line(size = 1) +
  labs(
    title = "Robbery Trends Over Time for States with Highest Population",
    x = "Year",
    y = "Total Robberies",
    color = "Jurisdiction"
  ) +
   scale_x_continuous(
    breaks = seq(min(top_jurisdiction_data$report_year), max(top_jurisdiction_data$report_year), by = 5)
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

```

```{r}
### Aggregate homicides data
jurisdiction_homicide_data <- data %>%
  group_by(jurisdiction = sub(",.*", "", agency_jurisdiction), report_year) %>%
  summarize(total_homicides = sum(homicides, na.rm = TRUE)) %>%
  ungroup()

### Create smaller set: jurisdictions with the highest total homicides across all years
top_jurisdiction <- jurisdiction_homicide_data %>%
  group_by(jurisdiction) %>%
  summarize(total_homicides_all_years = sum(total_homicides, na.rm = TRUE)) %>%
  arrange(desc(total_homicides_all_years)) %>%
  slice(1:10) %>%
  pull(jurisdiction)

### Exclude "United States"
top_jurisdiction <- setdiff(top_jurisdiction, "United States")

### Filter data for top jurisdictions and years 1975 and 2015
data_1975 <- jurisdiction_homicide_data %>%
  filter(jurisdiction %in% top_jurisdiction, report_year == 1975)

data_2015 <- jurisdiction_homicide_data %>%
  filter(jurisdiction %in% top_jurisdiction, report_year == 2015)

### Histogram for 1975
ggplot(data_1975, aes(x = jurisdiction, y = total_homicides, fill = jurisdiction)) +
  geom_col() +
  labs(
    title = "Total Homicides in 1975 for Top Jurisdictions",
    x = "Jurisdiction",
    y = "Total Homicides",
    fill = "Jurisdiction"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 10, angle = 45, hjust = 1),
    legend.position = "none"
  )

### Histogram for 2015
ggplot(data_2015, aes(x = jurisdiction, y = total_homicides, fill = jurisdiction)) +
  geom_col() +
  labs(
    title = "Total Homicides in 2015 for Top Jurisdictions",
    x = "Jurisdiction",
    y = "Total Homicides",
    fill = "Jurisdiction"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 10, angle = 45, hjust = 1),
    legend.position = "none"
  )


```

```{r}
### First Visualization: Homicides

# Aggregate homicides data
jurisdiction_homicide_data <- data %>%
  group_by(jurisdiction = sub(",.*", "", agency_jurisdiction), report_year) %>%
  summarize(total_homicides = sum(homicides, na.rm = TRUE)) %>%
  ungroup()

# Create smaller set: jurisdictions with the highest total homicides
top_jurisdiction <- jurisdiction_homicide_data %>%
  group_by(jurisdiction) %>%
  summarize(total_homicides_all_years = sum(total_homicides, na.rm = TRUE)) %>%
  arrange(desc(total_homicides_all_years)) %>%
  slice(1:10) %>%
  pull(jurisdiction)

# Filter data for top jurisdictions
top_jurisdiction_data <- jurisdiction_homicide_data %>%
  filter(jurisdiction %in% top_jurisdiction)

# Convert report_year to numeric for proper axis labeling
top_jurisdiction_data <- top_jurisdiction_data %>%
  mutate(report_year = as.numeric(as.character(report_year)))

# Line chart
ggplot(top_jurisdiction_data, aes(x = report_year, y = total_homicides, color = jurisdiction, group = jurisdiction)) +
  geom_line(size = 1) +
  labs(
    title = "Homicide Trends Over Time for Jurisdictions with Highest Homicides",
    x = "Year",
    y = "Total Homicides",
    color = "Jurisdiction"
  ) +
  scale_x_continuous(
    breaks = seq(min(top_jurisdiction_data$report_year), max(top_jurisdiction_data$report_year), by = 5)
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

```

```{r}
###Second visualization
# Filter data for 1975 and 2013 for top population
filtered_data <- data %>%
  filter(report_year %in% c(1975, 2013))

filtered_data <- filtered_data %>%
  mutate(homicide_rate = (homicides / population) * 100000)

# 10 jurisdictions by population in 2013
top_jurisdictions <- filtered_data %>%
  filter(report_year == 2013) %>%
  arrange(desc(population)) %>%
  slice(1:10) %>%
  pull(agency_jurisdiction)

# Filter data for only top 10 jurisdictions
top_data <- filtered_data %>%
  filter(agency_jurisdiction %in% top_jurisdictions)

# stacked bar chart
ggplot(top_data, aes(x = agency_jurisdiction, y = homicide_rate, fill = factor(report_year))) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Homicide Rates (1975 vs. 2013) for Top 10 Jurisdictions by Population",
    x = "Jurisdiction",
    y = "Homicide Rate (per 100,000)",
    fill = "Year"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(size = 10, hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10)
  )




```

```{r}
##Part 2 of second visualization
# Calculate homicide rate (per 100,000 population)
filtered_data <- filtered_data %>%
  mutate(homicide_rate = (homicides / population) * 100000)

# Find the 10 jurisdictions with the lowest population in 2013
lowest_jurisdictions <- filtered_data %>%
  filter(report_year == 2013) %>%
  arrange(population) %>%
  slice(1:10) %>%
  pull(agency_jurisdiction)

# Filter data for only these 10 jurisdictions
lowest_data <- filtered_data %>%
  filter(agency_jurisdiction %in% lowest_jurisdictions)

# Create the stacked bar chart
ggplot(lowest_data, aes(x = agency_jurisdiction, y = homicide_rate, fill = factor(report_year))) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Homicide Rates (1975 vs. 2013) for 10 Jurisdictions with Lowest Population",
    x = "Jurisdiction",
    y = "Homicide Rate (per 100,000)",
    fill = "Year"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(size = 10, hjust = 0.5),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

```

## flexdashboards

Do some research online about what flexdashboards are and what they look like. See if you can find some examples of existing dashboards and layouts. Describe here what single-page layout(s) you want to try.

Please make sure you include space for at least three plots and at least one descriptive panel.

**Important note**: You'll find some pretty neat interactive dashboards that incorporate shiny apps. We will not be covering that in this class, so please don't attempt it (...unless you really want to?)

I want to have more than 3 visualizations, I was thinking 4, and if I have time I want to try do an interactive map of the US with a homicide heat map. I want to have it in a square grid, with the description for the overall data as a bar on the top of the page. (Tabset row.)

What projects have you found that look nice/doable?

<https://testing-apps.shinyapps.io/flexdashboard-highcharter/>

I like this project because it has n example of a heat map for the US.

What layout(s) do you think you want to apply?

the Tabset row layout, but I don't want to just stick to that, I want to be able to be flexible depending on the amount of visualizations I need.

These visualizations are not final products, the one I am happy with is the line graph. I want to make all the visualizations the same aesthetic wise for my final dashboard.

```{r}
# Load required libraries
library(tidyverse)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(leaflet)
library(htmlwidgets)
library(usdata)

# Prepare crime data for 2015
data_2015 <- data %>%
  filter(report_year == 2015) %>%
  group_by(state = sub(".*, ", "", agency_jurisdiction)) %>%
  summarize(
    total_population = sum(population, na.rm = TRUE),
    total_violent_crimes = sum(violent_crimes, na.rm = TRUE),
    violent_crime_rate = (sum(violent_crimes, na.rm = TRUE) / sum(population, na.rm = TRUE)) * 100000
  ) %>%
  ungroup() %>%
  mutate(state = abbr2state(state)) # Convert abbreviations to full names

# Load U.S. state geometries
us_states <- ne_states(country = "United States of America", returnclass = "sf") %>%
  filter(!name %in% c("Hawaii", "Alaska")) # Exclude Hawaii and Alaska

# Merge crime data with spatial data
us_states <- us_states %>%
  left_join(data_2015, by = c("name" = "state"))

# Check for missing data
if (all(is.na(us_states$violent_crime_rate))) {
  stop("The violent_crime_rate column contains only NA values. Check the join or data preparation.")
}

# Create the heatmap with a title
map <- leaflet(us_states) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~colorNumeric(palette = "YlOrRd", domain = us_states$violent_crime_rate)(violent_crime_rate),
    weight = 1,
    color = "white",
    fillOpacity = 0.7,
    label = ~paste0(
      name, "<br>",
      "Violent Crime Rate: ", round(violent_crime_rate, 2), " per 100,000 people"
    )
  ) %>%
  addLegend(
    pal = colorNumeric(palette = "YlOrRd", domain = us_states$violent_crime_rate),
    values = ~violent_crime_rate,
    title = "Violent Crime Rate<br>(per 100,000)",
    position = "bottomright" # Move legend to top-left
  )

# Add a title to the map
map <- htmlwidgets::prependContent(
  map,
  htmltools::tags$h2("Violent Crime Rates US 2015", style = "text-align:center; margin: 10px;")
)

# Display the map
map

```

```{r}
#write.csv(jurisdiction_homicide_data, "C:/Users/faith/OneDrive/Documents/github/CS201_dashboard/data/jurisdiction_homicide_data.csv", row.names = FALSE)

write.csv(jurisdiction_robbery_data, "C:/Users/faith/OneDrive/Documents/github/CS201_dashboard/data/jurisdiction_robbery_data.csv", row.names = FALSE)

write.csv(data_2015, "C:/Users/faith/OneDrive/Documents/github/CS201_dashboard/data/data_2015.csv", row.names = FALSE)

```

```{r}
getwd()
```
