---
title: "EDA Assignment Prep"
author: "Amber Camp"
format: html
editor: visual
---

## EDA Assignment Prep: Pokemon data

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}
pokedata <- read_csv(here("data/Pokemon.csv"))

pokedata <-pokedata %>% 
  clean_names()

```

### Describe the data:

Data author: Alberto Barradas, via Kraggle.

<https://www.kaggle.com/datasets/abcsds/pokemon?resource=download>

Description: This is Pokemon data for 721 Pokemon and their stats.

How many observations (rows) are there? 800.

How many variables (columns) are there? 13.

### skimr: Quick overview

```{r}
skim_data <- skimr::skim(pokedata) #creted skim dataset
```

### Summary: Quick overview

```{r}
#summary(pokedata)

```

### Categorical vs. Continuous

```{r}
class(pokedata$number) 
pokedata$number <- as.factor(pokedata$number)
#summary(pokedata$number)

summary(pokedata$type_1) 
pokedata$type_1 <- as.factor(pokedata$type_1)
pokedata$generation <- as.factor(pokedata$generation) 
pokedata$type_2 <- as.factor(pokedata$type_2) 
```

How many categorical variables are there? 5-6.

How many continuous variables are there? 7.

What are some variables of interest to you? Speed, legendary, generation.

------------------------------------------------------------------------

## EDA Assignment Prep: \[Insert Data Name Here\]

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}

```

### Describe the data:

Data author:

Description:

How many observations (rows) are there?

How many variables (columns) are there?

### Summary: Quick overview

```{r}

```

### Categorical vs. Continuous

```{r}

```

How many categorical variables are there?

How many continuous variables are there?

What are some variables of interest to you?
